Water Management

### 2020-02-27 Paper/paper.tex
Minor overhaul, still some TODOs

In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
One of the most important uses of surface water is the irrigation of fields for food production. However, as farming and food production are private enterprises and lakes are public goods, interests of different parties have to be considered. 
The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities. 
Though looking at just two agents is a vast simplification compared to real allocation problems, one can easily add more stakeholders to the model, as long as there is enough computational power.
The sequence of events for every period is displayed in \ref{fig:Sequence}.
At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
The farmers utility function follows the form

@@

\subsubsection{Modification 2: Increasing the Volatility of Rainfalls}
In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility
We do so by setting $\sigma=2$ and $\mu=-1.5$.

@@

A common technique to solve optimization or combinatorial problems is called Dynamic Programming. 
It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). % Nicht eher "memoRization"?
In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
% TODO unbedingt noch Bellman als Quelle hinzufügen!!!
He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2). 
This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard to the state resulting from the first decisions" (\cite{bellman:1957}, p.2).

@@

So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$. 
%TODO hier noch auf die Definition von beta eingehen? Was ist beta und warum ist das relevant?
\subsection{Implementation}
%TODO theoretisch ist es sehr unschön ein Kapitel ohne Text aufzumachen. Ist aber nur Makulatur
\subsubsection{Computation of the value function}

@@

\subsection{Effect of the Discount Factor on the Optimal Irrigation Policy}
When altering the discount factor $\beta$ in a range from 0.05 to 0.95, the optimal irrigation policy changes. 
For small water levels of the reservoir, the optimal irrigation amount is rather insensitive to the discount factor. 
A small $\beta$ ensures that high water levels at the end of one period only keep their value only to a small extent in the next period.  
This leads to an increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much. 
A smaller $\beta$ allows more water to be used for irrigation today, optimizing the short-term utility. On the other hand, a higher $\beta$ increases the today's valuation of future utility, and therefore reduces the amount of water to be consumed (irrigated) today.
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/optimum-policy-over-beta.eps}
	\caption{Optimal irrigation policy as a function of the discount factor $\beta$}
	\label{fig:optimal-irrigation-beta}
\end{figure}
\newpage
\subsection{Increased Volatility of the Rainfalls} 
More volatile rainfalls means that rainfalls are having a higher standard deviation.
Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed. 
Originally, we assumed $\mu=0$ and $\sigma=1$. Now, we choose $\sigma=2$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).

\begin{equation}
	mean = exp(\mu+\frac{\sigma^2}{2}) \label{eq:rainfall}
\end{equation} 
Both distributions are visualized in \ref{fig:rain-distributions}.
\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{figures/rainfall-default.eps}
		\caption{$\mu=0, \sigma=1$}
		\label{fig:rainfall-default}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{figures/rainfall-more-volatile.eps}
		\caption{$\mu=-1.5, \sigma=2$}
		\label{fig:rainfall-changed}
	\end{subfigure}
	\caption{Different Rainfall Distributions}
	\label{fig:rain-distributions}
\end{figure}
\newpage
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-default.eps}
	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma=1)$}
	\label{fig:aggregated-use-rainfall-default}
\end{figure}
\begin{figure}[ht]
\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-more-volatile.eps}
\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma=2)$}
\label{fig:aggregated-use-rainfall-more-volatile}
\end{figure}

The following figures demonstrate the effect of the changed volatility. They show the lognormal rainfalls with the original volatility (\ref{fig:aggregated-use-rainfall-default}), and with the increased volatility (\ref{fig:aggregated-use-rainfall-more-volatile}) over 100 periods. 
When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit. 
%When comparing both visualizations,it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
%As a consequence the water level in the reservoir fluctuates much more leading to periods with almost no water in the reservoir. 
In the periods with only low precipitation, the farmers use almost no water for irrigation. 
The aggregated utility of farmers and recreational users drops significantly. 
Since the reservoir is capped at an amount of 7, the high peaks of the rainfalls can only partially be used in following (potentially dryer) periods. 
Generally speaking, one can say that a more volatile rain distribution leads to a lower average water level in the reservoir with lower amounts of water to be used for irrigation.
This effect is reflected in the aggregated utility as well. 
The utility of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$. 
Since the mean stays the same in the two cases considered, the value function does not change.

@@

With $\beta = 0.9$, the optimal irrigation amount tapers off again at the end, due to % TODO: Besprechen! MMn wegen Möglichkeit des Mehrfachnutzens bei Non-irrigation (daher nicht erkennbar bei beta = 0)
In figure 9 you can see one possible scenario in our basic setting with $\mu=0$ and $\sigma=1$.

@@

Increasing $\beta$ means putting more weight on future utility up to a factor of $\beta = 1$, where stakeholders are indifferent between current and future utility.

@@

The real rainfall distribution (\ref{fig:rainfall-real-data}) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma=1$.

@@

A whole different problem, more of an empirical nature however, would therefore be the determination of the "correct" discount factor, whatever that may be.
% TODO "whatever that may be" klingt echt merkwürdig
This is a different area of research and a lot of publications are dedicated to this topic but further advancements could greatly strengthen the theoretical foundation of our model.

>>>>

In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
One of the most important uses of surface water is the irrigation of fields for food production. However, as farming and food production are private enterprises and lakes are public goods, interests of different parties have to be considered. 
The water reservoir we're modeling has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities. 
Though looking at just two agents is a vast simplification compared to real allocation problems, one can easily add more stakeholders to the model, as long as there is enough computational power.
The sequence of events for every period is displayed in \ref{fig:Sequence}.
At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
The farmers utility function follows the form

@@

\subsubsection{Modification 2: Increasing the Volatility of Rainfalls}
In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility.
We do so by changing the parameters of the lognormal distribution of the rainfall to $\sigma^2=4$ and $\mu=-1.5$.

@@

A common technique to solve optimization or combinatorial problems is called Dynamic Programming. 
It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). % TODO Nicht eher "memoRization"?
In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
% TODO unbedingt noch Bellman als Quelle hinzufügen!!!
He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2). 
This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard to the state resulting from the first decisions" (\cite{bellman:1957}, p.2).

@@

So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$. 
\subsection{Implementation}
%TODO theoretisch ist es sehr unschön ein Kapitel ohne Text aufzumachen. Ist aber nur Makulatur
\subsubsection{Computation of the value function}

@@

\subsection{Effect of the Discount Factor on the Optimal Irrigation Policy}
When altering the discount factor $\beta$ in a range from 0.05 to 0.95, the optimal irrigation policy changes. 
For small water levels of the reservoir, the optimal irrigation amount is rather insensitive to the discount factor. 
A small $\beta$ ensures that high water levels at the end of one period only keep their value only to a small extent in the next period.  
This leads to an increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much. 
A smaller $\beta$ allows more water to be used for irrigation today, optimizing the short-term utility. On the other hand, a higher $\beta$ increases the today's valuation of future utility, and therefore reduces the amount of water to be consumed (irrigated) today.
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/optimum-policy-over-beta.eps}
	\caption{Optimal irrigation policy as a function of the discount factor $\beta$}
	\label{fig:optimal-irrigation-beta}
\end{figure}
\newpage
\subsection{Increased Volatility of the Rainfalls} 
More volatile rainfalls means that rainfalls are having a higher standard deviation.
Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed. % TODO: Das macht für mich nicht so ganz Sinn; entweder: "parameter mu has to also be changed..." oder "mean of the underlying normally distributed variable has to be changed"
Originally, we assumed $\mu=0$ and $\sigma^2=1$. Now, we choose $\sigma^2=4$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).

\begin{equation}
	mean = exp(\mu+\frac{\sigma^2}{2}) \label{eq:rainfall}
\end{equation} 
Both distributions are visualized in \ref{fig:rain-distributions}.
\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{figures/rainfall-default.eps}
		\caption{$\mu=0, \sigma^2=1$}
		\label{fig:rainfall-default}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{figures/rainfall-more-volatile.eps}
		\caption{$\mu=-1.5, \sigma^2=4$}
		\label{fig:rainfall-changed}
	\end{subfigure}
	\caption{Different Rainfall Distributions}
	\label{fig:rain-distributions}
\end{figure}
\newpage
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-default.eps}
	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma^2=1)$}
	\label{fig:aggregated-use-rainfall-default}
\end{figure}
\begin{figure}[ht]
\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-more-volatile.eps}
\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma^2=4)$}
\label{fig:aggregated-use-rainfall-more-volatile}
\end{figure}

The following figures demonstrate the effect of the changed volatility. They show the lognormal rainfalls with the original volatility (\ref{fig:aggregated-use-rainfall-default}), and with the increased volatility (\ref{fig:aggregated-use-rainfall-more-volatile}) over 100 periods. 
When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there are longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit. 
%When comparing both visualizations,it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
%As a consequence the water level in the reservoir fluctuates much more leading to periods with almost no water in the reservoir. 
In the periods with only low precipitation, the farmers use almost no water for irrigation. 
The aggregated utility of farmers and recreational users drops significantly. 
Since the reservoir is capped at an amount of 7, the high peaks of the rainfalls can only partially be used in following (potentially dryer) periods. 
Generally speaking, one can say that a more volatile rain distribution leads to a lower average water level in the reservoir with lower amounts of water to be used for irrigation.
This effect is reflected in the aggregated utility as well. 
The utility of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma^2=1$. 
Since the mean stays the same in the two cases considered, the value function does not change.

@@

With $\beta = 0.9$, the optimal irrigation amount tapers off again at the end, due to the potential of benefitting from the remaining water again in the future (see next paragraph).
In figure 9 you can see one possible scenario in our basic setting with $\mu=0$ and $\sigma^2=1$.

@@

Increasing $\beta$ means putting more weight on future utility in relation to current utility.
For example with a factor of $\beta = 1$, stakeholders are indifferent between current and future utility.

@@

The real rainfall distribution (\ref{fig:rainfall-real-data}) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma^2=1$.

@@

A whole different problem would therefore be the determination of the "correct" discount factor.
This is a different area of research and a lot of publications are dedicated to this topic. % TODO Quellen hinzufügen!
Still, further advancements could greatly strengthen the theoretical foundation of our model.
